---
title: "Assignment 5<br>Jericho  McLeod"
output: html_notebook
---

<b>Problem 1:</b> This problem concerns the automobile data from Assignments 3 and 4. As in Assignment 4, assume that counts of cars per 15-second interval are independent and identically distributed Poisson random variables with unknown mean L. Assume a uniform prior distribution for L. (This is a gamma distribution with shape 1 and scale infinity.) Using the posterior distribution from Problem 1 of Assignment 4, find the predictive distribution for the number of cars in the next 15-second interval. Name the family of distributions and the parameters of the predictive distribution. Find the predictive probability that 0, 1, 2, 3, 4, and more than 4 cars will pass the point in the next 15 seconds. Compare with your answer to Problem 1e of Assignment 3. Discuss.
```{r}
obs  <- c(0,0,0,1,1,1,1,1,2,2,2,2,2,2,2,3,3,3,4,4,4)
proba <- 1/length(obs)

alpha0=1; beta0=Inf               # prior shape and scale
alpha=alpha0+sum(obs)             # posterior shape
beta=1/(1/beta0+length(obs))      # posterior scale
y=0:4

for (i in 1:5) {
	fpred=dnbinom(y,size=alpha,prob=1/(1+1*beta))   # Negative binomial predictive distribution
	fpois=dpois(y,alpha*beta*1)                     # Poisson distribution for comparison
	PredPois=rbind(fpred,fpois)                     # Bind into a matrix
}

fpred  <- append(fpred,(1-sum(fpred))) # Append 5+ pred
fpois  <- c(0.148,0.277,0.264,0.172,0.086,0.053)
y <- append(y,'5+')

PredPois=rbind(fpred,fpois)  

barplot(PredPois,main=paste("Predictive Count of Cars"),
        xlab="Cars Arrived",
        ylab="Probability",
        names.arg=y,
        ylim=c(0,0.3),
        col=c("lightblue","blue"),
        beside=TRUE)
```
These are the same (excepting minor pixel count differences stemming from the number of significant digits). This is due the fact that the poisson distribution discrete distribution,  while the negative binomial is the likelihood of seeing  some number of successes before a failure. We have  essentially just used the negative binomial to reproduce the  probabilities obtained in the  poisson by using the shape and scale to find the probabililties of the discrete cuonts  from the poisson.  


<b>Problem 2:</b> In previous years, students in this course collected data on people’s preferences in the two Allais gambles from Assignment 2. For this problem, we will assume that responses are independent and identically distributed, and the probability is p that a person chooses both B in the first gamble and C in the second gamble.

A)  Assume that the prior distribution for $\pi$ is Beta(1, 3). Find the prior mean and standard deviation for $\pi$. Find a 95% symmetric tail area credible interval for the prior probability that a person would choose B and C. Do you think this is a reasonable prior distribution to use for this problem? Why or why not?
```{r}
Beta  <- c(1,3)
prior_mean <- (Beta[1] * 1) / (Beta[1]+Beta[2])
prior_var <- (1*Beta[1]*Beta[2]*(Beta[1]+Beta[2]+1))/((Beta[1]+Beta[2])^2 * (Beta[1]+Beta[2]+1))
prior_sd  <- prior_var^0.5

cred_int <- qbeta(c(0.025,0.975),shape1=Beta[1],shape2=Beta[2])

print(paste0("Prior Mean:  ",prior_mean))
print(paste0("Prior Standard Deviation:  ",prior_sd))
print("Prior Credible Interval:")
print(cred_int)
```
Beginning from a level distribution is acceptable without any expert opinion. In this case, we know that individuals do not typically choose purely at  random, but have bias towards answers. However, given that  we do not currently have any quantifiable expert opinions regarding  how much bias, a level distribution is reasonable.


B) In 2009, 19 out of 47 respondents chose B and C. Find the posterior distribution for the probability $\pi$ that a person in this population would choose B and C. Find the posterior mean and standard deviation, and a 95% symmetric tail area credible interval for $\pi$. Do a triplot.
```{r}
Beta_posterior <- Beta
Beta_posterior[1] <-  Beta[1] + 19
Beta_posterior[2] <- Beta[2] + (47-19)

posterior_mean <- (Beta_posterior[1] * 1) / (Beta_posterior[1]+Beta_posterior[2])
posterior_var <- (1*Beta_posterior[1]*Beta_posterior[2]*(Beta_posterior[1]+Beta_posterior[2]+1))/((Beta_posterior[1]+Beta_posterior[2])^2 * (Beta_posterior[1]+Beta_posterior[2]+1))
posterior_sd  <- posterior_var^0.5

cred_int <- qbeta(c(0.025,0.975),shape1=Beta[1],shape2=Beta[2])


Beta_uni <-c(1,1)
Norm_beta_post <- c(Beta_uni[1]+19,Beta_uni[2]+(47-19))

lambda <- seq(0,1,0.01)
prior_dens <- dbeta(lambda,shape1=Beta[1],shape2=Beta[2])
posterior_dens <-  dbeta(lambda,shape1=Beta_posterior[1],shape2=Beta_posterior[2])
norm_plot <- dbeta(lambda,shape1=Norm_beta_post[1],shape=Norm_beta_post[2])


plot(lambda,prior_dens,
     type="l",
     col="blue",
     main="Triplot for Allais Gamble",
     xlab="Choosing B and C",
     ylab="Prob Density",
     xlim=c(0,1),
     ylim=c(0,6))
lines(lambda,norm_plot,col='green')
lines(lambda,posterior_dens,col="red")
legend(0.7,5.0,c("Prior","Norm Lik","Posterior"),col=c("blue","green","red"),lty=c(1,1,1))


```


C) Find the predictive distribution for the number of B and C responses in a future sample of 50 people drawn from the same population. Compare with a Binomial distribution using a point estimate of the probability of choosing B and C.
```{r}
library("rmutil")

p = Beta_posterior[1]  / (Beta_posterior[1]+Beta_posterior[2])
m = Beta_posterior[1]+Beta_posterior[2]
n = 50

# Predictive Dist
pred_dist <-  dbetabinom(seq(0,50,1),n,p,m)
plot(seq(0,50,1),pred_dist,type="l",ylim=c(0,.15),col='green',     xlab="Trials",
     ylab="Probability  ",)

# Binomial  dist with point est
binom_dist <- dbinom(seq(0,50,1), size=50, prob=(19/47)) 
lines(seq(0,50,1),binom_dist,col='blue')

legend(30,0.15,c("Beta Binom Pred","Binomial"),col=c("blue","green","red"),lty=c(1,1,1))
```
D) Comment on your results.

The binomial point estimate  is more dispersed compared to the  beta binomial predictions  for number of successes in the next  50 trials. This is because, while the binomial distribution is the count of successes at a given probability, the beta binomial draws successes from a beta distribution. 